---
title: "documentation_api"
author: "Solichatus Zahroh"
date: "2023-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Start the REST API

### Docker Setup

1. Set up Docker Desktop by installing it from this web https://docs.docker.com/get-docker/. Docker Desktop requires a minimum of 2 GB of RAM and 20 GB of disk space to run.
2. In Windows, you might need to install WSL2. Open the command prompt and type `wsl --install` or if you already installed it, check for the updates `wsl --update`.
3. Keep open your Docker Desktop when trying to run below code. Keep in mind that the Dockerfile is located in the same path as your working directory.
```
docker build -t webmice -f docker/Dockerfile .
docker run -it -p 8080:8080 -v  C:/Users/Zenius/Documents/GitHub/webmice:/home/webmice/testdata webmice
# get a shell inside of the docker image for debugging
docker exec -it webmice bash
```
Change `</path/to/local/datafolder>`to your local path. You only need the `-v </path/to/local/datafolder>:/home/webmice/testdata` flag if you want to make some of your local data known to the docker image to use them as input data for the HTTP API. 

### Testing the Swagger
You can test to upload your dataset in Swagger API. The variables must be less than 50 variables (number of columns) and in CSV format.

##### **`/version`**: Obtain version information

Let us first check whether WEBMICE is running. The following code makes a simple request to WEBMICE to see whether it is alive and to return the version number of the underlying `mice` R package. We illustrate both requests in `R` and in `bash`.

We first need to install and load packages.

```{r httr, eval=FALSE}
#install.packages(c("httr2", "jsonlite"))
```
```{r load}
library(httr2)
library(jsonlite)
```

Let's find out the WEBMICE version number. In this document, we define the server that hosts WEBMICE as follows, and optionally create a variable `.host` for use in the bash terminal.

```{r host}
host <- "http://localhost:8080"
```

We first illustrate a method that makes a request to the server. The following command prepares a GET request `req` to the `/version` end point:

```{r req_version}
req <- request(host) |> 
  req_headers(accept = "text/plain") |> 
  req_url_path("/version")
req
```

The next step sends the request to the API server, and stores the result in the server response `resp`:

```{r call_version}
resp <- req_perform(req)
resp
```

The next step parses the `resp` object and converts the body into an R list:

```{r ret_version}
ret <- resp |> 
  resp_body_string() |> 
  fromJSON()
ret
```

See the documentation of the `httr2` package for other operations and extractions. One may chain the above operations as:

```{r}
request(host) |> 
  req_headers(accept = "text/plain") |> 
  req_url_path("/version") |> 
  req_perform() |> 
  resp_body_string() |> 
  fromJSON()
```

##### **`/data`**: Upload data

Uploading data to WEBMICE can be done by sending a text file. Suppose that `../testdata/tempdata.csv` is a comma-delimited file with data. We may upload it as
```{r upload_r, eval = TRUE}
req <- request(host) |> 
  req_url_path("/data") |> 
  req_headers(accept = "text/plain", 
              `content-type` = "multipart/form-data") |>
  req_body_multipart(csvfile = curl::form_file("C:/Users/Zenius/Documents/GitHub/webmice/testdata/nhanes.csv"))
resp <- req_perform(req, verbosity = 3)
token <- resp$headers$data_token
resp_headers(resp)
resp_header_exists(resp, "data_token")
```

The file is uploaded to the server. The `data_token` header may be used for subsequent requests to refer to the uploaded data.

##### **`/long`**: Impute missing data 
`imp <- mice(nhanes, maxit = 2, m = 2, seed = 1)`

Impute missing data from the uploaded file.

```{r upload_r, eval = TRUE}
library(RCurl)
# Create the payload object
payload <- list(data = "nhanes", maxit = 2, m = 2, seed = 1)

# Encode the payload object to JSON
payload_json <- jsonlite::toJSON(payload, auto_unbox =  TRUE)

# Create the full URL
full_url <- paste(host, "/long","?payload=", curlEscape((payload_json)), sep = "")

long <- request("full_url") |> 
  req_headers(accept = "text/plain", 
              `content-type` = "multipart/form-data") |>
  req_body_multipart(input = resp$headers$data_token)
resp_long <- req_perform(long, verbosity = 3) |>
  resp_body_string() |> 
  fromJSON()
```

Change the `data` to `data_token` from the uploaded data.

##### **`/fit`**: Fit the model
`fit <- with(data = imp, exp = lm(chl ~ age + bmi))`
```{r upload_r, eval = TRUE}
library(RCurl)
# Create the payload object
payload_fit <- list(data =resp_long$result, model = lm, formula = chl~age+bmi){"data": [{".imp":1,".id":1,"age":1,"bmi":30.1,"hyp":1,"chl":187}], "model":["lm"], "formula":["chl ~ age + bmi"]}

# Encode the payload object to JSON
payload_json_fit <- jsonlite::toJSON(payload_fit, auto_unbox =  TRUE)

# Create the full URL
full_url_fit <- paste(host, "/fit","?payload=", curlEscape((payload_json_fit)), sep = "")

fit <- request("full_url_fit") |> 
  req_headers(accept = "text/plain", 
              `content-type` = "multipart/form-data") |>
  req_body_multipart(input = resp$headers$data_token)
resp_fit <- req_perform(fit, verbosity = 3) |>
  resp_body_string() |> 
  fromJSON()
```
Get the fitted values of the imputed file.

##### **`/pool`**: Pool the results of fitting a model to the imputed data. 
`summary((fit))`

Get summary of the fitted model.

```{r upload_r, eval = TRUE}
library(RCurl)
# Create the payload object
payload_fit <- list(data =resp_fit$result, model = lm, formula = chl~age+bmi){"data": [{".imp":1,".id":1,"age":1,"bmi":30.1,"hyp":1,"chl":187}], "model":["lm"], "formula":["chl ~ age + bmi"]}

# Encode the payload object to JSON
payload_json_fit <- jsonlite::toJSON(payload_fit, auto_unbox =  TRUE)

# Create the full URL
full_url_fit <- paste(host, "/fit","?payload=", curlEscape((payload_json_fit)), sep = "")

fit <- request("full_url_fit") |> 
  req_headers(accept = "text/plain", 
              `content-type` = "multipart/form-data") |>
  req_body_multipart(input = resp$headers$data_token)
resp_fit <- req_perform(fit, verbosity = 3) |>
  resp_body_string() |> 
  fromJSON()
```


### Get the Version of the API
```{r}
# Get version
library(httr)
host <- "http://localhost:8080/"
path <- "version"
url <- parse_url(host)
url <- modify_url(url, path = file.path(url$path, path))
r <- GET(url)
fromJSON(content(r, type = "text", encoding = "UTF-8"))
```
