---
title: WEBMICE - 
subtitle: webmice 0.0.1 (December 2023)
author: 
output:
  html_document:
  number_sections: false
  self-contained: true
theme: united
editor_options: 
  chunk_output_type: console
---
<!-- README.md is generated from README.Rmd. Please edit that file -->
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
```

## Overview

WEBMICE is a web service for imputing and analysing incomplete datasets. WEBMICE provides a language-independent API to the functionality of the MICE R package. 
It can 

1. read data from local system of the client;
2. create multiple complete versions from incomplete data;
3. analyse each of the imputed datasets;
4. pool the analyses results into one result.

WEBMICE is a RESTful API that runs on a (remote) host. In principle, any `HTTP` client will work with WEBMICE. The following sections illustrate how a client can make requests to WEBMICE using various client languages. 

### Primary WEBMICE end points

  | Verb  | API Endpoint          | Description                                | Maps to `mice` function |
  |:------|:-------------------------- |:------------------------------------------ |:-------------------------|
  | GET  | `/version`       | Get the `mice`'s version                          | `utils::packageVersion("mice")`          |
  | GET  | `/exampledata`       | Get example of an incomplete data (`nhanes` and `nhanes2`)            | `mice::nhanes`  or `mice::nhanes2`   |
  | POST  | `/data`       | Upload an incomplete data in CSV format                          |          |
  | GET  | `/long`       | Impute missing data            | `mice::mice()`    
  |
  | GET  | `/fit`       | Get some predefined fit value                          | `mice::getfit(with())`        |
  | GET  | `/pool`       | Pool results of a fit given its summary table            | `mice::pool()`    
  |
The table lists the defined API end points and the internal mapping between the API end point and the corresponding R function. 

### Objective

This document provides a quick introduction into the main WEBMICE features, and how these can be assessed from `R` and from the bash command line.


## Features 

### {.tabset}

#### **R**

##### **`/version`**: Obtain version information

Let us first check whether WEBMICE is running. The following code makes a simple request to WEBMICE to see whether it is alive and to return the version number of the underlying `mice` R package. We illustrate both requests in `R` and in `bash`.

We first need to install and load packages.

```{r httr, eval=FALSE}
#install.packages(c("httr2", "jsonlite"))
```
```{r load}
library(httr2)
library(jsonlite)
```

Let's find out the WEBMICE version number. In this document, we define the server that hosts WEBMICE as follows, and optionally create a variable `.host` for use in the bash terminal.

```{r host}
host <- "http://localhost:8080"
```

We first illustrate a method that makes a request to the server. The following command prepares a GET request `req` to the `/version` end point:

```{r req_version}
req <- request(host) |> 
  req_headers(accept = "text/plain") |> 
  req_url_path("/version")
req
```

The next step sends the request to the API server, and stores the result in the server response `resp`:

```{r call_version}
resp <- req_perform(req)
resp
```

The next step parses the `resp` object and converts the body into an R list:

```{r ret_version}
ret <- resp |> 
  resp_body_string() |> 
  fromJSON()
ret
```

See the documentation of the `httr2` package for other operations and extractions. One may chain the above operations as:

```{r}
request(host) |> 
  req_headers(accept = "text/plain") |> 
  req_url_path("/version") |> 
  req_perform() |> 
  resp_body_string() |> 
  fromJSON()
```

##### **`/exampledata`**: Impute built-in dataset
```{r}
req_ex <- request(host) |> 
  req_headers(accept = "text/json") |> 
  req_url_path("/exampledata") |> 
  req_url_query(name = "nhanes")
req_ex |> req_dry_run()
resp_ex <- req_perform(req_ex)
ret_ex <- resp_ex |> 
  resp_body_string() |> 
  fromJSON()
head(ret_ex$result)
```

##### **`/long`**: Impute built-in dataset

Raw call

```{r}
resp <- request("http://localhost:8080/long?payload=%7B%22data%22%3A%22nhanes%22%2C%22maxit%22%3A2%2C%22m%22%3A2%2C%22seed%22%3A1%7D") %>% 
  req_method("GET") %>% 
  req_headers(accept = "text/json") %>% 
  req_perform()
ret <- resp |> 
  resp_body_string() |> 
  fromJSON()
head(ret$result)
```

A more user-friendly call
```{r, eval=TRUE}

req_long <- request(host) |> 
  req_headers("accept" = "text/json") |> 
  req_url_path("/long") |> 
  req_url_query(payload = '{"data":"nhanes","maxit":2,"m":1,"seed":1}')
req_long |> req_dry_run()
resp_long <- req_perform(req_long)
ret_long <- resp_long |> 
  resp_body_string() |> 
  fromJSON()
head(ret_long$result)

```

##### **`/fit`**: Some predifined fits

All result from `/long` as input for `/fit`. A more user-friendly call (data comes from `complete(..., action = 'long', include = TRUE)`)

```{r, eval=TRUE}
body_long <- (toJSON(ret_long$result))

req_fit <- request(host) |> 
  req_headers("accept" = "text/json") |>
  req_url_path("/fit") |> 
  req_url_query(payload = paste0('{"data":', body_long,', "model":["lm"], "formula":["chl ~ age + bmi"]}'))
req_dry_run(req_fit)
resp_fit <- req_fit |> req_perform()
ret_fit <- resp_fit |> 
  resp_body_string() |> 
  fromJSON()
ret_fit$result
```


##### **`/pool`**: Pool results of a fit given its summary table (mice 3.14.4 or higher)
```{r}
resp2 <- request("http://localhost:8080/pool?payload=%7B%22data%22%3A%20%5B%7B%22term%22%3A%22%28Intercept%29%22%2C%22estimate%22%3A23.9357%2C%22std.error%22%3A3.8732%2C%20%22statistic%22%3A6.1798%2C%22p.value%22%3A3.2106e-06%2C%22nobs%22%3A25%2C%22df.residual%22%3A22%7D%2C%20%7B%22term%22%3A%22hyp%22%2C%22estimate%22%3A1.1677%2C%22std.error%22%3A1.6218%2C%22statistic%22%3A0.72%2C%20%22p.value%22%3A0.4791%2C%22nobs%22%3A25%2C%22df.residual%22%3A22%7D%2C%7B%22term%22%3A%22chl%22%2C%22estimate%22%3A0.0063%2C%20%22std.error%22%3A0.0185%2C%22statistic%22%3A0.3385%2C%22p.value%22%3A0.7382%2C%22nobs%22%3A25%2C%22df.residual%22%3A22%7D%2C%20%7B%22term%22%3A%22%28Intercept%29%22%2C%22estimate%22%3A24.5434%2C%22std.error%22%3A4.7611%2C%22statistic%22%3A5.155%2C%20%22p.value%22%3A0%2C%22nobs%22%3A25%2C%22df.residual%22%3A22%7D%2C%7B%22term%22%3A%22hyp%22%2C%22estimate%22%3A1.5843%2C%20%22std.error%22%3A2.236%2C%22statistic%22%3A0.7086%2C%22p.value%22%3A0.486%2C%22nobs%22%3A25%2C%22df.residual%22%3A22%7D%2C%20%7B%22term%22%3A%22chl%22%2C%22estimate%22%3A-0.0032%2C%22std.error%22%3A0.0205%2C%22statistic%22%3A-0.1569%2C%20%22p.value%22%3A0.8768%2C%22nobs%22%3A25%2C%22df.residual%22%3A25%7D%5D%7D") %>% 
  req_method("GET") %>% 
  req_headers(accept = "text/json") %>% 
  req_perform()
ret2 <- resp2 |> 
  resp_body_string() |> 
  fromJSON()
head(ret2$result)

```

A more user-friendly call

```{r, eval=TRUE}

body_fit<- as.character(toJSON(ret_fit$result))

req_pool <- request(host) |> 
  req_headers("accept" = "text/json") |>
  req_url_path("/pool") |> 
  req_url_query(payload = paste0('{"data":', body_fit,'}')) 
req_dry_run(req_pool)
resp_pool <- req_pool |> req_perform()
ret_pool <- resp_pool |> 
  resp_body_string() |> 
  fromJSON()
ret_pool$result
```


##### **`/data`**: Upload data

Uploading data to WEBMICE can be done by sending a text file. Suppose that `../testdata/tempdata.csv` is a comma-delimited file with data. We may upload it as

```{r upload_r, eval = TRUE}
input <- curl::form_file("C:/Users/Zenius/Documents/GitHub/webmice/testdata/tempdata.csv", type = "text/csv")
#input <- curl::form_file("testdata/tempdata.csv", type = "text/csv")
resp <- request(host) |> 
  req_url_path("/data") |> 
  req_headers("accept" = "text/plain", 
              "content-type" = "multipart/form-data") |>
  req_body_multipart(csvfile = input) |> 
  req_perform()
resp_header(resp, "data_token")
```

The file is uploaded to the server. The `data_token` header may be used for subsequent requests to refer to the uploaded data.

##### **`/long`**: Impute data using the token (not yet working)

```{r impute_r, eval=TRUE}
token <- resp_header(resp, "data_token")
token_with_quotes <- paste0('"', token, '"')
print(token_with_quotes)

req_longdata <- request(host) |> 
  req_headers("accept" = "text/json") |>
  req_url_path("/long") |> 
  req_url_query(payload = paste0('{"data":', token_with_quotes,  ',"maxit":2,"m":1,"seed":1}')) 
req_longdata |> req_dry_run()
resp_longdata <- req_longdata |> req_perform()
ret_longdata <- resp_longdata |> 
  resp_body_string() |> 
  fromJSON()
head(ret_longdata$result)
```

##### **`/fit`**: Some predifined fits

```{r, eval=TRUE}
body_longdata <- as.character(toJSON(ret_longdata$result))

req_fitdata <- request(host) |> 
  req_headers("accept" = "text/json") |>
  req_url_path("/fit") |> 
  req_url_query(payload = paste0('{"data":', body_longdata,', "model":["lm"], "formula":["chl ~ age + bmi"]}'))
req_dry_run(req_fitdata)
resp_fitdata <- req_fitdata |> req_perform()
ret_fitdata <- resp_fitdata |> 
  resp_body_string() |> 
  fromJSON()
ret_fitdata$result

```

##### **`/pool`**: Pool results of a fit given its summary table (mice 3.14.4 or higher)

```{r, eval=TRUE}

body_fitdata<- as.character(toJSON(ret_fitdata$result))

req_pooldata <- request(host) |> 
  req_headers("accept" = "text/json") |>
  req_url_path("/pool") |> 
  req_url_query(payload = paste0('{"data":',body_fit,'}') )
req_dry_run(req_pooldata)
resp_pooldata <- req_pooldata |> req_perform()
ret_pooldata <- resp_pooldata |> 
  resp_body_string() |> 
  fromJSON()
ret_pooldata$result
```

#### **Python**

```{python}
import requests
from urllib.parse import urljoin
import urllib.parse
import json
import re

# Specify the url for host
host = 'http://localhost:8080'
```

##### **`/version`**: Obtain version information

```{python}
## /version

# Add path
path_version = '/version'

# Combine the url
version = urljoin(host,path_version)

# Add headers
headers = {'Accept': 'application/json'}

# Get the url
response_version = requests.get(version, headers=headers)

# See the response
response_version.content
```

##### **`/exampledata`**: Get the example data

There are two different example data sets, namely 'nhanes' and 'nhanes2'
```{python}
# /exampledata (nhanes)

# Add path
endpoint = "/exampledata"

# Construct the query parameters
params = {'name': 'nhanes'}

# Additional headers if needed
headers = {'Accept': 'application/json'}

# Make the GET request
response = requests.get(host + endpoint, params=params, headers=headers)

# Print the response
print(response.text)
```

##### **`/long`**: Impute built-in dataset
```{python}
## /long

# Add path
path_long = '/long'

# Combine the url
long = urljoin(host,path_long)

# Add headers
headers = {'Accept': 'application/json'}

# Add payload
params = {'data':'nhanes','maxit':2,'m':1,'seed':1}

# Encode the payload as a JSON string and then URL-encode the entire payload
payload = urllib.parse.quote('{"data":"%s","maxit":%d,"m":%d,"seed":%d}' % (params['data'], params['maxit'], params['m'], params['seed']))

# Construct the complete URL
long_url = f"{long}?payload={payload}"

# Make a request using the complete URL
response = requests.get(long_url)

# Get the response body
response.json()
json_response = response.json()

# Access the "result" part
result_part = json_response.get("result", [])
result_part

```

##### **`/fit`**: Some predifined fits
``` {python}
## /fit

# Add path
path_fit = '/fit'

# Combine the url
fit = urljoin(host,path_fit)

# Add headers
headers = {'Accept': 'application/json'}

# Add payload
params_fit = {
    "data": result_part,
    "model": ["lm"],
    "formula": ["chl ~ age + bmi"]
}
# Encode the payload as a JSON string and then URL-encode the entire payload
payload_fit = json.dumps(params_fit)
url_encoded_payload_fit = urllib.parse.quote(payload_fit)

# Construct the complete URL
fit_url = f"{fit}?payload={url_encoded_payload_fit}"

# Make a request using the complete URL
response_fit = requests.get(fit_url)

# Get the response body
json_response_fit = response_fit.json()

# Access the "result" part
result_fit = json_response_fit.get("result", [])
result_fit
```

##### **`/pool`**: Pool results of a fit given its summary table (mice 3.14.4 or higher)
```{python}
## /pool

# Add path
path_pool = '/pool'

# Combine the url
pool = urljoin(host,path_pool)

# Add headers
headers = {'Accept': 'application/json'}

# Add payload
params_pool = {"data": result_fit}

# Encode the payload as a JSON string and then URL-encode the entire payload
payload_pool = json.dumps(params_pool)
url_encoded_payload_pool = urllib.parse.quote(payload_pool)

# Construct the complete URL
pool_url = f"{pool}?payload={url_encoded_payload_pool}"

# Make a request using the complete URL
response_pool = requests.get(pool_url)

# Get the response body
json_response_pool = response_pool.json()

# Access the "result" part
result_pool = json_response_pool.get("result", [])
result_pool
```

##### **`/data`**: Upload data

Uploading data to WEBMICE can be done by sending a text file. Suppose that `../testdata/tempdata.csv` is a comma-delimited file with data. We may upload it as
```{python}
# /data

# Add  path
path_data = '/data'

# Combine the url
data_upload = urljoin(host,path_data)

# File path of the csv file
file_path = r"C:\Users\Zenius\Documents\GitHub\webmice\testdata\tempdata.csv"

# Construct the files dictionary
files = {'csvfile': (file_path, open(file_path, 'rb'))}

# Additional headers
headers = {'accept': 'text/plain'}

# Make the POST request
response = requests.post(data_upload, files=files, headers=headers)

# Print all response headers
for key, value in response.headers.items():
    print(f"{key}: {value}")

# Retrieve the data_token from headers
data_token = response.headers.get('data_token')
print(data_token)
```

##### **`/long`**: Impute built-in dataset

This part is exactly the same procedure as the previous one using example data (nhanes). The only difference is the data part for the payload. We use `params = {'data':data_token,'maxit':2,'m':1,'seed':1}` instead of `params = {'data':'nhanes,'maxit':2,'m':1,'seed':1}`

```{python}
## /long

# Add path
path_long = '/long'

# Combine the url
long = urljoin(host,path_long)

# Add headers
headers = {'Accept': 'application/json'}

# Add payload
params = {'data':data_token,'maxit':2,'m':1,'seed':1}

# Encode the payload as a JSON string and then URL-encode the entire payload
payload = urllib.parse.quote('{"data":"%s","maxit":%d,"m":%d,"seed":%d}' % (params['data'], params['maxit'], params['m'], params['seed']))

# Construct the complete URL
long_url = f"{long}?payload={payload}"

# Make a request using the complete URL
response = requests.get(long_url)

# Get the response body
response.json()
json_response = response.json()

# Access the "result" part
result_part = json_response.get("result", [])
result_part

```

For `/fit` and `/pool` parts, follow the same steps using the example data.
##### **`/fit`**: Some predifined fits
``` {python}
## /fit

# Add path
path_fit = '/fit'

# Combine the url
fit = urljoin(host,path_fit)

# Add headers
headers = {'Accept': 'application/json'}

# Add payload
params_fit = {
    "data": result_part,
    "model": ["lm"],
    "formula": ["chl ~ age + bmi"]
}
# Encode the payload as a JSON string and then URL-encode the entire payload
payload_fit = json.dumps(params_fit)
url_encoded_payload_fit = urllib.parse.quote(payload_fit)

# Construct the complete URL
fit_url = f"{fit}?payload={url_encoded_payload_fit}"

# Make a request using the complete URL
response_fit = requests.get(fit_url)

# Get the response body
json_response_fit = response_fit.json()

# Access the "result" part
result_fit = json_response_fit.get("result", [])
result_fit
```

##### **`/pool`**: Pool results of a fit given its summary table (mice 3.14.4 or higher)
```{python}
## /pool

# Add path
path_pool = '/pool'

# Combine the url
pool = urljoin(host,path_pool)

# Add headers
headers = {'Accept': 'application/json'}

# Add payload
params_pool = {"data": result_fit}

# Encode the payload as a JSON string and then URL-encode the entire payload
payload_pool = json.dumps(params_pool)
url_encoded_payload_pool = urllib.parse.quote(payload_pool)

# Construct the complete URL
pool_url = f"{pool}?payload={url_encoded_payload_pool}"

# Make a request using the complete URL
response_pool = requests.get(pool_url)

# Get the response body
json_response_pool = response_pool.json()

# Access the "result" part
result_pool = json_response_pool.get("result", [])
result_pool
```


#### **bash**
We use the `curl` Linux command. If needed, on Ubuntu install `curl` as

```{bash, eval=FALSE}
sudo apt update
sudo apt -y install curl
```


##### **`/version`**: Obtain version information
Let’s find out the WEBMICE version number. We first illustrate a method that makes two requests to the server.

The following `bash` commands call the `/version` API end point with a GET request, and stores the response in file `resp`. 

The response to the request consists of a set of URLs created on the server, each of which contains details on the response. 

```{bash, eval = FALSE}
echo http://localhost:8080 > .host

curl \
 -X GET "$(cat .host)/version" \
 -H "accept: text/plain" \
 -o resp \

cat resp
```

##### **`/exampledata`**: Obtain example data
The R package `mice` contains a few built-in datasets (See <https://amices.org/mice/reference/index.html#datasets>). These datasets can be imputed and analysed directly on the server. This is to get example data of `nhanes`

```{bash, eval = FALSE}
curl \
  -X GET "$(cat .host)/exampledata?name=nhanes" \
  -H "accept: text/json" \
  -o nhanes.json \

cat nhanes.json
```

##### **`/data`**: Upload data
Uploading data to WEBMICE can be done by sending a text file.

Suppose that `../testdata/tempdata.csv` is a comma-delimited file with data. We may upload it as:
(This part should be change to your full local path `F "csvfile=@fullpath/tempdata.csv;type=text/csv"`).

From this part, we can obtain data_token from the uploaded csv file

```{bash, eval = FALSE}
curl -i \
  -X POST "$(cat .host)/data" \
  -H "accept: text/plain" \
  -H "Content-Type: multipart/form-data" \
  -F "csvfile=@C:/Users/Zenius/Documents/GitHub/webmice/testdata/tempdata.csv;type=text/csv" \
  -o response.txt \

cat response.txt

# Get the data_token from headers
data_token=$(grep -i '^data_token:' response.txt | awk '{print $2}')
echo "$data_token"
```
The file is uploaded to the server. The `data_token` header may be used for subsequent requests to refer to the uploaded data.

##### **`/long`**: Obtain version information

Get the long data for `nhanes` data
```{bash, eval = FALSE}
curl -G "$(cat .host)/long" \
  -H "accept: application/json" \
  --data-urlencode  'payload={"data":"nhanes","maxit":2,"m":1,"seed":1}' \
  -o outputlong.json \

cat outputlong.json
```

Get the long data for uploaded data using the data_token
```{bash, eval = FALSE}
curl -G -v "$(cat .host)/long" \
  -H "accept: application/json" \
  --data-urlencode 'payload={"data":"'$data_token'","maxit":2,"m":1,"seed":1}' \
  -o longdata.json \

cat longdata.json
```


##### **`/fit`**: Obtain version information
```{bash, eval = FALSE}
# Extract the result from JSON and remove spaces
result_long=$(jq -r '.result' outputlong.json | tr -d '[:space:]')

# Encode the result_long as JSON string and remove spaces
result_long=$(echo "$result_long" | jq -c -r)

# Construct the payload with the encoded result_long
payload="{\"data\":$result_long,\"model\":[\"lm\"],\"formula\":[\"chl~age+bmi\"]}"

# Use the payload in the curl command
curl -G -v "$(cat .host)/fit" \
  -H "accept: application/json" \
  --data-urlencode "payload=$payload" \
  -o outputfit.json

```

##### **`/pool`**: Obtain version information

```{bash, eval = FALSE}
# Extract the result from JSON and remove spaces
result_fit=$(jq -r '.result' outputfit.json | tr -d '[:space:]')

# Encode the result_long as JSON string and remove spaces
result_fit=$(echo "$result_fit" | jq -c -r)

# Construct the payload with the encoded result_long
payload="{\"data\":$result_fit}"

# Use the payload in the curl command
curl -G -v "$(cat .host)/pool" \
  -H "accept: application/json" \
  --data-urlencode "payload=$payload" \
  -o outputpool.json
```

### {-}

## Resources

### Internal

| Description                                           | Status                                  |
|:------------------------------------------------------|:----------------------------------------|

### External
  
| Description                                           | Status                                  |
|:------------------------------------------------------|:----------------------------------------|
